---
title: "BSMM-lab-3"
subtitle: "BSMM 8740 Fall 2023"
author: "Meghana Ghanta"
date: "October 2nd, 2023"
format: html
editor: visual
self-contained: true
---

## Setup

```{r}
library(magrittr)   # the pipe
library(tidyverse)  # for data wrangling + visualization
library(tidymodels) # for modeling
library(gt)         # for pretty tables?
```

```{r}
#install.packages("skimr")
library(skimr)

```

```{r}
#install.packages("janitor")
library(janitor)
```

```{r}
library(recipes)
library(dplyr)
```

Load packages and data:

```{r load-pkg-data}
#| message: false
boston_cocktails <- readr::read_csv('data/boston_cocktails.csv')
```

## Exercises

### Exercise 1

```{r}
skimr::skim(boston_cocktails)
DataExplorer::introduce(boston_cocktails)
  
```

```{r}
summary(boston_cocktails$measure_number, category="Cocktail Classics")
```

The median measure amount across across all cocktails is 1.

### Exercise 2

```{r}
boston_cocktails1 <- boston_cocktails |>
  dplyr::select(name, category, ingredient, measure_number)
pivot_cocktails <- boston_cocktails1 |>
  pivot_wider(
    names_from =ingredient,
    values_from=measure_number)

pivot_cocktails <-pivot_cocktails |>
  replace(is.na(pivot_cocktails),0)

pivot_cocktails <- pivot_cocktails |>
  clean_names()
```

```{r}
gin_amount <- pivot_cocktails |>
  dplyr::filter(name=="Leap Frog Highball")
gin_amount
```

The **Leap Frog Highball** contains 2 of gin

### Exercise 3

```{r}
rec <- recipe( ~., data = boston_cocktails) |>
  step_mutate(name, category, role = "id") |>
  step_select(where(is.numeric)) |>
  step_normalize(all_predictors()) |>
  step_pca(all_predictors())

prepped_data <- prep(rec)

num_predictors <- length(names(prepped_data)) - 2  
num_predictors
```

7 predictor variables are prepped by the recipe.

### Exercise 4

```{r}

step_info <- tidy(prepped_data)

normalization_info <- tidy(prepped_data, 1)

ingredient_counts1 <- boston_cocktails1 |>
  dplyr::group_by(ingredient) |>
  dplyr::summarise(average_measure = mean(measure_number, na.rm = TRUE)) |>
  dplyr::arrange(desc(average_measure))

most_used_ingredient <- ingredient_counts1$ingredient[1]
most_used_ingredient
```

On average the most used ingredient in the Boston Cocktails dataset is cranberry juice.

### Exercise 5

```{r}


```

Describe describe the drinks represented by PC1?

### Exercise 6

```{r}

```

The characteristic alcoholic beverage of each of the first 4 principle components is \_\_\_.

### Exercise 7

How would you interpret the results of a PCA analysis for a client?
